\addcontentsline{toc}{chapter}{Acknowledgments}
\thispagestyle{empty}

\vspace*{20mm}

\begin{center}
{\usekomafont{section} Acknowledgments}
\end{center}

\vspace{10mm}

Before anything else, I would like to express my utmost gratitude to my advisor, Han Liang. Thank you for guiding me with patience and diligence throughout the entire process. It is only thanks to her indispensable help that I am able to complete this work and I am extremely grateful for this. I would also like to say thank you to my supervisor, Prof. Dr. Rüdiger Westermann, for granting me the opportunity to pursue this topic and affording me this invaluable experience.

Additionally, I would like to thank all the parties that have indirectly been involved in the creation of this work--thank you to Louis Bidou for creating the immaculate \textit{Chocolate Bunny}, and also to creators of the depth extraction models involved in this study. Thank you for laying down the key foundations to my work and thank you for your contributions in the 3D Graphics community.

To you, the readers of my work: thank you for being a witness to the past few years of my life. I hope that in these pages you will find something useful to aid you in your own journeys.

With that said, I would also like to express my gratitude to my cherished friends and family. I know it has been a long time coming, but thank you for your unwavering support and belief in me. Through the highs and lows of life, you are a constant glow in my world and I'm forever grateful for your presence.

Finally, I would like to say thank you to my special person. Thank you for the warm hugs and the endless nights you spent accompanying me. You always make my days full of color and happiness. As I open the door to my next journey in life, I hope you will always be here to hold my hands.

\cleardoublepage{}
